Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
use subcaption instead of obsolete package subfigure
added subsubsection: "Example: 2-D Jacobi iteration"

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

\subsubsection{Example: 2-D Jacobi iteration}

One relatively basic example of a stencil program is two dimensional
Jacobi iteration that repeatedly goes through each cell in a matrix
and computes the average value of the four adjacent cells. The kernel
can be expressed in a line of Fortran code,
\begin{minted}{fortran}
  a(i,j) = (a(i-1,j)+a(i+1,j)+a(i,j+1)+a(i,j-1))/4
\end{minted}
and Camfort is able to infer a precise stencil specification for this
line,
\begin{minted}{fortran}
  != stencil reflexive(dim=1)*centered(depth=1, dim=2, irreflexive) + reflexive(dim=2)*centered(depth=1, dim=1, irreflexive) :: a
\end{minted}
which when inserted back into the program can be checked by Camfort as
well. We then tested the error-detecting ability of Camfort by
generating a family of 6,561 Jacobi iteration programs by perturbing
the indices used in the following manner: each access of the array was
templated as \mintinline{fortran}{a(i+ki,j+kj)} and then all possible
combinations of \mintinline{fortran}{ki} and \mintinline{fortran}{kj}
were drawn from the set $\{-1,0,1\}$. There are four uses of
\mintinline{fortran}{ki} and four of \mintinline{fortran}{kj}, hence
$3^8=6561$ possible perturbations of this kind. This particular setup
tests the class of errors in which the programmer typed an incorrect
number but at least got the ordering of the indices correct, although
Camfort would also detect a mixup of variables in this case as well.

Camfort checked all 6,561 programs and found that only 24 of them were
valid implementations of the stencil specification. This was the
expected result because there are $4!$ possible correct permutations
of the four terms in the kernel. All of the other programs were found
to violate the specification in some way. Another way to test the
programs is through visual examination of their output in graphical
format. Figure~\ref{fig:jacobi} shows some of the results of running
the Jacobi iteration programs, rendered into image files. All of the
correct variants emitted images indistinguishable from
Figure~\ref{fig:jacobicorrect}, while all of the incorrect variants
had wildly different image outputs. Both the ``visual test'' and a
byte-by-byte comparison showed results that completely concurred with
the results of the Camfort test.

\begin{figure}[t]
  \centering
  \subcaptionbox{\label{fig:jacobicorrect}}{\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  \subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  \subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  \subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \caption{Greyscale output of several perturbations of a Jacobi
    iteration program. The correct output is shown in
    (\subref{fig:jacobicorrect}).}
  \label{fig:jacobi}
\end{figure}
